ESI for "Probabilistic determination of the effect of a deep eutectic solvent on the structure of lipid monolayers"

### 2018-07-23 reports/paper.tex
second draft

\includegraphics{head_foot/DOI} & \noindent\LARGE{\textbf{Probabilistic determination of the effect of a deep-eutectic solvent on the structure of lipid monolayers}} \\%Article title goes here instead of the text "This is the title"
\vspace{0.3cm} & \vspace{0.3cm} \\

 & \noindent\large{Adrian Sanchez-Fernandez,\textit{$^{ab}$}$^{\dag\P}$ Andrew R. McCluskey,\textit{$^{ac}$}$^{\dag}$ Karen J. Edler,\textit{$^{a}$}$^{\ast}$ Andrew J. Jackson,\textit{$^{bd}$} Richard A. Campbell,\textit{$^{e}$} and Thomas Arnold\textit{$^{abcf}$}$^{\ast}$} \\%Author names go here instead of "Full name", etc.

\includegraphics{head_foot/dates} & \noindent\normalsize{Deep eutectic solvents present a novel class of non-aqueous room temperature solvent, with tunable properties and capable of promoting self-assembly of surfactant molecules. However, the solvation model in these systems still challenges the classic understanding of amphiphilicity. In this work, we present the first example of the self-assembly of a phospholipid monolayer at the interface between air and a non-aqueous solvent. Furthermore, we use novel, chemically-relevant modelling of reflectometry measurements to show the ability for the deep eutectic solvent to have an interaction with the phosphocholine lipid head group, leading to an apparent reduction in the component volume compared to that observed in water. No such reduction was observed for the phosphatidylglycerol head group, indicating that the interaction is ion specific.} \\

@@

\footnotetext{\textit{$^{a}$~Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK.}}
\footnotetext{\textit{$^{b}$~European Spallation Source, SE-211 00 Lund, Sweden.}}
\footnotetext{\textit{$^{c}$~Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK.}}
\footnotetext{\textit{$^{d}$~Department of Physical Chemistry, Lund University, SE-211 00 Lund, Sweden.}}
\footnotetext{\textit{$^{e}$~Institut Laue-Langevin, 71 avenue des Martyrs, 38000, Grenoble, France.}}
\footnotetext{\textit{$^{f}$~ISIS Neutron and Muon Source, Science and Technology Facilities Council, Rutherford Appleton Laboratory, Harwell Oxford, Didcot OX11 0QX, UK.}}
\footnotetext{\textit{$^{\P}$~Present address: Department of Food Technology, Lund University, SE-211 00 Lund, Sweden}}
\footnotetext{$^{\ast}$~Corresponding author: k.edler@bath.ac.uk, tom.arnold@esss.se}

@@

%%%MAIN TEXT%%%%
\section{Introduction}
Deep eutectic solvents (DES) are green, sustainable solvents obtained through the complexation of naturally occurring compounds, such as sugars, alcohols, amines and carboxylic acids, amoung others.\cite{Smith2014, Dai2013} An extensive hydrogen-bonding network is present between these precursors, allowing the mixture to remain liquid at room temperature due to the high-entropic state of the mixture.\cite{Hammond2016, Hammond2017, Araujo2017} Additionally, through different combinations of the precursors materials, it is possible to tune the physicochemical properties of the solvent, such as polarity,\cite{Pandey2014} viscosity and surface tension,\cite{Smith2014} network charge,\cite{Zahn2016} and hydrophobicity.\cite{Ribeiro2015,vanOsch2015}

These solvents have recently shown the ability to promote the self-assembly of surfactants into micellar structures\cite{Sanchez-Fernandez2016,Arnold2015} and to stabilise the conformation of non-ionic polymer species,\cite{Sapir2016} indicating the presense of a solvophobic effect. The behaviour and conformation of biomolecules in DES has seen and increase in interest,\cite{Esquembre2013,Gorke2010,Gorke2008,Monhami2014,Wu2014,Harifi-Mood2017,Milano2017,Sanchez-Fernandez2017} due to potential applications in the preservation of biomolecules, as environments for enzymatic reactions.\cite{Merza2018} Futhermore, recent investigations have also shown that DES have been shown to support the formation of phospholipid bilayers.\cite{Bryant2017,Bryant2016,Gutierrez2009}

@@

Recent developments in computational resource and software has enabled powerful methodologies and algorithms to be harnessed by those from non-computer science backgrounds. This has mostly occurred through open-source software projects such as the Python language and the Jupyter notebooks framework.\cite{vanRossum1995,Kluyver2016} In the area of neutron and X-ray reflectometry data-analysis, this has led to the development of refnx,\cite{Nelson2018} a Python library for the fitting of layer-based models to reflectometry data. refnx enables the use of custom models that contain chemically-relevant information. This information includes constaints such as that the number density of phospholipid headgroups must be the same as the number density of pairs of phospholipid tailgroups and that the length of the tail chains should not surpass a maximum of the Tanford length for a given chain.\cite{Tanford1980}

The use of a Python library for fitting enables powerful probability distribution function (PDF) sampling methods to be used such as the Goodman \& Weare Affine Invariant Markov chain Monte Carlo (MCMC) Ensemble,\cite{Goodman2010} as implemented in the Python library emcee.\cite{Foreman-Mackey2013} This method allows for the sampling of a high-dimensionality parameter space, such as that which is relevant to reflectometry fitting, to be performed with relative ease. This results in estimatations of the inverse uncertainties associated with each parameter as well as information about the correlations between different parameters within the model.

In this work, we present the first investigation of the structure of phospholipid monolayers at the air-DES interface, as determined by chemical-context modelling of X-ray reflectometry (XRR) measurements. Four different phospholipids; 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (DMPG), were studied at the interface between a 1:2 mixture of choline chloride:glycerol and air. This allowed the nature of two, chemically distinct, phospholipid headgroups to be understood in this non-aqueous solvent, in addition to the effect of the tail chain length.

@@

\subsection{Materials}
Choline chloride (99 \%, Sigma-Aldrich) and glycerol (99 \%, Sigma-Aldrich) d$_9$-choline chloride (99 \%, 98 \% D, CK Isotopes) and d8-glycerol (99 \%, 98 \% D, CK Isotopes)  were purchased and used without further purification. The DES was prepared by mixing the precursors at the appropriate mole ratio, and heating at 80 $^\circ$C until a homogeneous, transparent liquid formed.\cite{Smith2014} The solvent was equilibrated overnight at 40 $^\circ$C and subsequently stored under a dry atmosphere. Due to the limited availability of the deuterated precursors, a fully protonated subphase (hDES) and a partially deuterated subphase (hdDES) were prepared and used during the neutron reflectometry (NR) experiment. The partiall deuterated subpahse was prepared using the following mixtures of precursors: 1 mole of 0.38 fraction of h-choline chloride/0.62 mole fraction of d-choline chloride; and 2 moles of 0.56 mole fraction of h-glycerol/0.44 mole fraction of d-glycerol. The solvent was subsequently prepared follow the procedure discussed above.

The water content on the DES was determined before and after each experiment by Karl-Fischer titration (Mettler Toledo DL32 Karl-Fischer Coulometer, Aqualine Electrolyte A, Aqualine Catholyte CG A) in order to ensure water presence was kept to a minimum. Those measurements showed that the water content of the solvent was kept below 0.3 wt\% during all the experimental procedures presented here, which we assume to be negligible and have little impact on the characteristics of the DES.\cite{Hammond2016,Hammond2017}

DPPC (> 99 \%), DMPC (> 99 \%), and DMPG (> 99 \%) were supplied by Avanti Polar Lipids and DLPC (> 99 \%) was supplied by Sigma-Aldrich and all were used as received. Deuterated versions of DPPC (d62-DPPC, > 99 \%, deuterated tails-only) and DMPC (d54-DPPC, > 99 \%, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification. These phospholipids were dissolved in chloroform (0.5 mg/mL) at room temperature.

In the XRR, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration. After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface. Surface concentration was modified by closing and opening the PTFE barriers of the Langmuir trough. In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a managageable level) it was not possible to use a Langmuir trough. Instead, small Delrin adsorption troughs were used that did not have controlable barriers. This resulted in no control over the surface pressure of the measurement and therefore it was not possible to co-refine different NR contrasts.

\subsection{Methods}
XRR measurements were taken on I07 at Diamond Light Source, at 12.5 keV photon energy using the double-crystal-deflector.\cite{Arnold2012} The reflected intensity was measured in a momentum transfer range from 0.018 to 0.7 \AA$^{-1}$. The data was normalised with respect to the incident beam and the background was measured from off-specular reflection and subsequently subtracted. The sample environment consisted of a PTFE Langmuir trough. Samples were equilibrated for at least one hour and preserved under argon atmosphere to minimise the adsorption of water by the subphase. Surface concentration was controlled through opening and closing the barriers of the Langmuir trough. XRR data were collected for each of the lipids, DMPC, DPPC, DLPC and DMPG at two surface pressures, 20 mNm$^{-1}$ and 30 mNm$^{-1}$, as measured with an aluminium Wilhelmy plate, all measurements were made at 22 $^\circ$C.

The NR experiments were performed on FIGARO at the Institut Laue-Langevin using the time-of-flight method.\cite{Campbell2011} Two angle regimes were measured to provide a momentum transfer range from 0.0005 to 0.18 \AA$^{-1}$. Small-volume Delrin adsorption troughs were used to support the sample. A single surface concentration for each system and contrast was measured. Similar to the X-ray procedure, samples were given enough time to equilibrate (at least two hours) and kept under inert atmosphere.

\subsection{Data analysis}
The use of reflectometry to analyse the structure of phospholipids on the surface of water has a histroy extending over many years.\cite{Mohwald1990,Kewalramani2010,Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990} This has led to some variation in the models used to fit experimental data, shown in Table \ref{tab:water}. Although, there appears to be a general consensus that the component volume of the phosphocholine is in the range from 320 \AA$^3$ to 360 \AA$^3$, and as a result this volume is often used as a physical constraint on the layer model when fitting reflectometry data. However, since this work involves a non-aqueous solvent we do not know whether the head group component volumes used in the literature are derived from water-based measurements are appropriate for this work. The charged nature of the zwitterionic and anionic lipid heads means that they are likely to have different interactions with neutral water as compared to the charged DES.\cite{Sanchez-Fernandez2018}

@@

To allow for the use of a chemically-relevant model, where the lipid head group component volume was allowed to vary, the Python library refnx\cite{Nelson2018} as used. This software allows the inclusion of a custom chemically-relevant model from which the parameters to be fed into the Abel\`{e}s layer-model,\cite{Abeles1950,Parratt1954} that is typical for reflectometry fitting, are obtained. This custom model, along with Jupyter notebook showing in full the analysis performed, can be found in the ESI.

This chemically-relevant model involves considering the lipids as consisting of head groups and tail groups. The head groups have a calculated scattering length, $b_h$, (found as a summation of the number of electrons in the head group multiplied by the classical radius of the electron), and a component volume, $V_h$. These head groups make up a layer with a given thickness, $d_h$, and roughness, $\sigma_h$, within which some volume fraction of solvent can intercalate, $\phi_h$. The tail groups also have a calculated scattering length, $b_t$, and a component volume, $V_t$, however the thickness of the tail group layer, $d_t$, is found from the length of the carbon tail, $t_t$, and angle that the chain is tilted by with respect to the interface normal, $\theta_t$, 
\begin{equation}
\label{equ:tl}
d_t = t_t \cos{\theta_t}.
\end{equation}
The air above the monolayer is also able to solvate the tail region by some volume fraction, $\phi_t$. The scattering length density (SLD) of the tail and head layers used in the Abel\`{e}s model can therefore be found as follows, 
\begin{equation}
\text{SLD}_i = \frac{b_i}{V_i}(1 - \phi_i) + \text{SLD}_{s}\phi_i,
\end{equation}
where $\text{SLD}_{s}$ is the scattering length density of the superphase (air) and the subphase (DES) for the tail and head groups respectively, and $i$ indicates either the tail or head layer. In order to ensure that the number density of head groups and pairs of tail groups is the same, the following constraint was added to the model,\cite{Braun2017}
\begin{equation}
\label{equ:ht}
d_h = \frac{V_hd_t(1-\phi_t)}{V_t(1-\phi_h)}. 
\end{equation}

Initially, this custom model was used to co-refine the component volume of the lipid head group, $V_h$, for the two surface concentration XRR measurements. Some parameters were allowed to vary ($\theta_t$, $\phi_t$, $\phi_h$, $\sigma_t$, and $\sigma_h$); while others, shown in Table \ref{tab:invariant}, were held constant at the values given. The length of the carbon chain was kept constant to the value determined by the Tanford equation,\cite{Tanford1980} this is valid due to the condensed nature of the monolayer at this surface concentration resulting in the extended, staggered conformation of the chain being likely.

\begin{table}[h]
	\small
	\caption{\ The invariant parameters within the chemically-sensible model. $^a$Values determined from Armen \emph{et al.}\cite{Armen1998} $^b$Values obtained from the Tanford formula where the carbon chains are assumed to be fully extended.\cite{Tanford1980} $^c$Values extracted from Sanchez-Fernandez \emph{et al.}\cite{Sanchez-Fernandez2016}}
	\label{tab:invariant}
	\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}llllll}
		\hline
		Component & $b_t$/fm & $b_h$/fm & $V_t$/\AA$^3$ & $t_t$/\AA & $\text{SLD}$/$\times10^{-6}$\AA$^{-2}$ \\
		\hline
		X-ray & & & & & \\
		DLPC & 5073 & 4674 & 667$^a$ & 15.5$^b$ & -- \\
		DMPC & 5985 & 4674 & 779$^a$ & 18.0$^b$ & -- \\
		DPPC & 6897 & 4674 & 891$^a$ & 20.5$^b$ & -- \\
		DMPG & 5985 & 4731 & 779$^a$ & 18.0$^b$ & --\\
		Air & -- & -- & -- & -- & 0\\
		DES & -- & -- & -- & -- & 10.8$^c$ \\
		\hline
		Neutron & & & & & \\
		d54-DMPC & 5329.8 & 602.7 & 779$^a$ & 18.0$^b$ & -- \\
		d62-DPPC & 6129.2 & 602.7 & 891$^a$ & 20.5$^b$ & -- \\
		h-DES & -- & -- & -- & -- & 0.43$^c$  \\
		hd-DES & -- & -- & -- & -- & 3.15$^c$ \\
		\hline
	\end{tabular*}
\end{table}

The determined head group component volumes were then used in the refinement of the custom model against the NR measurements. Due to the lack of contrast present between the hydrogenated phosphocholine head group and the solvent, it was necessary to constrain the thickness and roughness of the head layer based on that determined from the highest surface concentration XRR measurement, this allowed the chain tilt angle to be constrained based on the following relationship, 
\begin{equation}
\label{equ:ct}
\theta_t = \cos^{-1}\bigg(\frac{V_td_h(1-\phi_h)}{V_ht_t(1-\phi_t)}\bigg),
\end{equation}
ensuring that the number density of head groups and pairs of tails is held constant. Table \ref{tab:invariant} gives the details of the scattering length and SLDs used as invariant parameters in the custom model, while the tail component volumes and tail chain length were the same as for the XRR.

The refinement of the custom model to the experimental data involved transforming the reflectometry calculated from the model and the data into $Rq^4$ such that the artifact of the Fresnel decay was removed, before using the differential evoulation method available to refnx from the scipy library,\cite{Jones2001} to find the parameters that gave the best fit to the data. The parameter space was then probed using the MCMC method available through emcee, this allowed for an estimate of the PDF associated with each parameter. In the MCMC sampling, 200 walkers were used over 1000 iterations, following an equilibration of 200 iterations.

@@

\subsection{Solvent effect on lipid head component volumes}

>>>>

\includegraphics{head_foot/DOI} & \noindent\LARGE{\textbf{Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers}} \\%Article title goes here instead of the text "This is the title"
\vspace{0.3cm} & \vspace{0.3cm} \\

 & \noindent\large{Andrew R. McCluskey,\textit{$^{ac}$}$^{\dag}$ Adrian Sanchez-Fernandez,\textit{$^{ab}$}$^{\dag\P}$ Karen J. Edler,\textit{$^{a}$}$^{\ast}$ Andrew J. Jackson,\textit{$^{bd}$} Richard A. Campbell,\textit{$^{ef}$} and Thomas Arnold\textit{$^{abcg}$}$^{\ast}$} \\%Author names go here instead of "Full name", etc.

\includegraphics{head_foot/dates} & \noindent\normalsize{Deep eutectic solvents present a novel class of non-aqueous room temperature solvent with tunable properties, that are capable of promoting self-assembly of surfactant molecules. However, the solvation model in these systems still challenges the classic understanding of amphiphilicity. In this work, we present the first example of the self-assembly of phospholipid monolayers at the interface between air and a non-aqueous solvent. Furthermore, we use novel, chemically-consistant Bayesian modelling of reflectometry measurements to show the ability of the deep eutectic solvent to interact with the phosphatidylglycerol lipid head group, leading to an apparent increase in the component volume compared to that observed in water. No such change was observed for the phosphocholine head group, indicating that the interaction is head group specific.} \\

@@

\footnotetext{\textit{$^{a}$~Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK.}}
\footnotetext{\textit{$^{b}$~European Spallation Source, SE-211 00 Lund, Sweden.}}
\footnotetext{\textit{$^{c}$~Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK.}}
\footnotetext{\textit{$^{d}$~Department of Physical Chemistry, Lund University, SE-211 00 Lund, Sweden.}}
\footnotetext{\textit{$^{e}$~Institut Laue-Langevin, 71 avenue des Martyrs, 38000, Grenoble, France.}}
\footnotetext{\textit{$^{f}$~Division of Pharmacy and Optometry, University of Manchester, Manchester, M13 9PT, UK.}}
\footnotetext{\textit{$^{g}$~ISIS Neutron and Muon Source, Science and Technology Facilities Council, Rutherford Appleton Laboratory, Harwell Oxford, Didcot OX11 0QX, UK.}}
\footnotetext{\textit{$^{\P}$~Present address: Department of Food Technology, Lund University, SE-211 00 Lund, Sweden}}
\footnotetext{$^{\ast}$~Corresponding author: k.edler@bath.ac.uk, tom.arnold@esss.se}

@@

%%%MAIN TEXT%%%%
\section{Introduction}
Deep eutectic solvents (DES) are green, sustainable solvents obtained through the complexation of naturally occurring compounds, such as sugars, alcohols, amines and carboxylic acids, among others.\cite{Smith2014, Dai2013} An extensive hydrogen-bonding network is present between these precursors, allowing the mixture to remain liquid at room temperature due to its high-entropic state.\cite{Hammond2016, Hammond2017, Araujo2017} Additionally, through different combinations of the precursors materials, it is possible to tune the physicochemical properties of the solvent, such as polarity,\cite{Pandey2014} viscosity and surface tension,\cite{Smith2014} network charge,\cite{Zahn2016} and hydrophobicity.\cite{Ribeiro2015,vanOsch2015}

These solvents have recently shown the ability to promote the self-assembly of surfactants into micellar structures\cite{Sanchez-Fernandez2016,Arnold2015} and to stabilise the conformation of non-ionic polymer species,\cite{Sapir2016} indicating the presense of a solvophobic effect. The behaviour and conformation of biomolecules in DES have seen an increase in interest,\cite{Esquembre2013,Gorke2010,Gorke2008,Monhami2014,Wu2014,Harifi-Mood2017,Milano2017,Sanchez-Fernandez2017} due to potential applications in the preservation of biomolecules, as environments for enzymatic reactions.\cite{Merza2018} Futhermore, recent investigations have also shown that DES have been able to support the formation of phospholipid bilayers.\cite{Bryant2017,Bryant2016,Gutierrez2009}

@@

Recent developments in computational resource and software have enabled powerful methodologies and algorithms to be harnessed by those from non-expert backgrounds. This has mostly occurred through open-source software projects such as the Python language and the Jupyter notebooks framework.\cite{vanRossum1995,Kluyver2016} In the area of neutron and X-ray reflectometry data-analysis, this has led to the development of refnx,\cite{Nelson2018} a Python library for the fitting of layer-based models to reflectometry data. The refnx library enables the use of custom models that contain chemically-relevant information. This information includes constaints such as that the number density of phospholipid headgroups must be the same as the number density of pairs of phospholipid tailgroups and that the length of the tail chains should not surpass a maximum of the Tanford length for a given chain.\cite{Tanford1980} This length is an emperically calculated value for the maximum expansion of a series of carbon atoms.

The use of a Python library for fitting enables powerful probability distribution function (PDF) sampling methods to be used such as the Goodman \& Weare Affine Invariant Markov chain Monte Carlo (MCMC) Ensemble,\cite{Goodman2010} as implemented in the Python library emcee.\cite{Foreman-Mackey2013} This method allows for the Bayesian sampling of a high-dimensionality parameter space, such as that which is relevant to reflectometry fitting, to be performed with relative ease. This results in estimations of the inverse uncertainties associated with each parameter as well as information about the correlations between different parameters within the model.

In this work, we present the first investigation of the structure of phospholipid monolayers at the air-DES interface, as determined by chemically-consistant modelling of X-ray reflectometry (XRR) measurements. Four different phospholipids; 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (DMPG), were studied at the interface between a 1:2 mixture of choline chloride:glycerol and air. This allowed the nature of two, chemically distinct, phospholipid headgroups to be understood in this non-aqueous solvent, in addition to the effect of the tail chain length. The analysis was then extended to model complementary neutron reflectometry measurements for two contrasts of DMPC and DPPC at a single surface pressure.

@@

\subsection{Materials}
Choline chloride (99 \%, Sigma-Aldrich) and glycerol (99 \%, Sigma-Aldrich) d$_9$-choline chloride (99 \%, 98 \% D, CK Isotopes) and d$_8$-glycerol (99 \%, 98 \% D, CK Isotopes)  were purchased and used without further purification. The DES was prepared by mixing the precursors at the appropriate mole ratio, and heating at 80 $^\circ$C until a homogeneous, transparent liquid formed.\cite{Smith2014} The solvent was equilibrated overnight at 40 $^\circ$C and subsequently stored under a dry atmosphere. Due to the limited availability of the deuterated precursors, a fully protonated subphase (hDES) and a partially deuterated subphase (hdDES) were prepared and used during the neutron reflectometry (NR) experiment. The partially deuterated subphase was prepared using the following mixtures of precursors: 1 mole of 0.38 fraction of h-choline chloride/0.62 mole fraction of d-choline chloride; and 2 moles of 0.56 mole fraction of h-glycerol/0.44 mole fraction of d-glycerol. The solvent was subsequently prepared following the procedure discussed above.

The water content of the DES was determined before and after each experiment by Karl-Fischer titration (Mettler Toledo DL32 Karl-Fischer Coulometer, Aqualine Electrolyte A, Aqualine Catholyte CG A) in order to ensure water presence was kept to a minimum. Those measurements showed that the water content of the solvent was kept below 0.3 wt\% during all the experimental procedures presented here, which we assume to be negligible and have little impact on the characteristics of the DES.\cite{Hammond2016,Hammond2017}

DPPC (> 99 \%, C$_{16}$ tails), DMPC (> 99 \%, C$_{14}$ tails), and DMPG (> 99 \%, C$_{14}$ tails) were supplied by Avanti Polar Lipids and DLPC (> 99 \%, C$_{12}$ tails) was supplied by Sigma-Aldrich and all were used as received. Deuterated versions of DPPC (d$_{62}$-DPPC, > 99 \%, deuterated tails-only) and DMPC (d$_{54}$-DPPC, > 99 \%, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification. These phospholipids were dissolved in chloroform (0.5 mg/mL) at room temperature.

In the XRR experiment, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration. After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface. Surface concentration was modified by closing and opening the PTFE barriers of a Langmuir trough. In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a managageable level) it was not possible to use a Langmuir trough. Instead, small Delrin adsorption troughs were used that did not have controlable barriers. This resulted in no control over the surface pressure of the measurement and therefore it was not possible to co-refine different NR contrasts.

\subsection{Methods}
XRR measurements were taken on I07 at Diamond Light Source, at 12.5 keV photon energy using the double-crystal-deflector.\cite{Arnold2012} The reflected intensity was measured in a momentum transfer range from 0.018 to 0.7 \AA$^{-1}$. The data were normalised with respect to the incident beam and the background was measured from off-specular reflection and subsequently subtracted. Samples were equilibrated for at least one hour and preserved under an argon atmosphere to minimise the adsorption of water by the subphase. XRR data were collected for each of the lipids, DMPC, DPPC, DLPC and DMPG at two surface pressures, 20 mNm$^{-1}$ and 30 mNm$^{-1}$, as measured with an aluminium Wilhelmy plate; all measurements were made at 22 $^\circ$C.

The NR experiments were performed on FIGARO at the Institut Laue-Langevin using the time-of-flight method.\cite{Campbell2011} Data at two incident angles of 0.62$^\circ$ and 3.8$^\circ$ were measured to provide a momentum transfer range from 0.005 to 0.18 \AA$^{-1}$. A single surface concentration for each system and contrast was measured. Similar to the X-ray procedure, samples were given enough time to equilibrate (at least two hours) and kept under an inert atmosphere.

\subsection{Data analysis}
The use of reflectometry to analyse the structure of phospholipids on the surface of water has a history extending over many years.\cite{Mohwald1990,Kewalramani2010,Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990,Campbell2018} This has led to some variation in the models used to fit experimental data, shown in Table \ref{tab:water}. There appears to be a general consensus that the component volume of the phosphocholine (PC) headgroup is in the range from 320 \AA$^3$ to 360 \AA$^3$ while data from a single source gives the phosphatidylglycerol (PG) headgroup to be 291 \AA$^3$, and as a result these volumes is often used as physical constraints on the layer model when fitting reflectometry data. However, since the current work involves a non-aqueous solvent we do not know whether the head group component volumes used in the literature that are derived from water-based measurements will be appropriate for this work. The charged nature of the zwitterionic and anionic lipid heads means that they are likely to have different interactions with neutral water as compared to the charged DES.\cite{Sanchez-Fernandez2018} Further it has been shown that in the liquid condensed phase, which is present at high surface pressure, the lipid tails will become compressed resulting in a component volume less than measured in the liquid expanded phase.\cite{Campbell2018}

@@

To allow for the use of a chemically-consistant model, where the lipid component volumes were allowed to vary, the Python library refnx\cite{Nelson2018} as used. This software allows the inclusion of a custom model from which the parameters to be fed into the Abel\`{e}s model for the reflection of light at a given number of stratified interfaces,\cite{Abeles1950,Parratt1954} that is typical for reflectometry fitting, are obtained. This custom model, along with a series of Jupyter notebooks showing, in full, the analysis performed, can be found in the ESI and is available under an MIT license.\cite{mccluskey_2018}

This chemically-consistent model involves two layers consisting of head groups at the interface with the solvent and tail groups at the interface with the air. The head groups have a calculated scattering length, $b_h$, (found as a summation of the number of electrons in the head group multiplied by the classical radius of the electron), and a component volume, $V_h$. These head groups make up a layer with a given thickness, $d_h$, and roughness, $\sigma_h$, within which some volume fraction of solvent can intercalate, $\phi_h$. The tail groups also have a calculated scattering length, $b_t$, and a component volume, $V_t$, however the thickness of the tail group layer, $d_t$, is found from the length of the carbon tail, $t_t$, and angle that the chain is tilted by with respect to the interface normal, $\theta_t$, 
\begin{equation}
\label{equ:tl}
d_t = t_t \cos{\theta_t}.
\end{equation}
The scattering length density (SLD) of the tail and head layers used in the Abel\`{e}s model can therefore be found as follows, 
\begin{equation}
\text{SLD}_i = \frac{b_i}{V_i}(1 - \phi_i) + \text{SLD}_{s}\phi_i,
\end{equation}
where $\text{SLD}_{s}$ is the scattering length density of the subphase (DES) for the tail and head groups respectively, and $i$ indicates either the tail or head layer, not that it is assumed that no solvent penetrates into the tail layer, e.g. $\phi_t = 0$. To ensure that the number density of head groups and pairs of tail groups is the same, the following constraint was added to the model,\cite{Braun2017}
\begin{equation}
\label{equ:ht}
d_h = \frac{V_hd_t(1-\phi_t)}{V_t(1-\phi_h)}. 
\end{equation}
A single value for the interfacial roughness was fitted for all interfaces, as there is only a single lipid molecule type in each monolayer.

In the first of two steps, this custom model was used to co-refine the component volume of the lipid head group, $V_h$, and the volume of the tail group, $V_t$, across the two surface concentration XRR measurements. The following parameters were allowed to vary; $\theta_t$, $d_h$, $\phi_h$, and $\sigma_{t,h,s}$, independently across the two surface pressurse. While others, shown in Table \ref{tab:invariant}, were held constant at the values given. The length of the carbon chain was kept constant to the value determined by the Tanford equation,\cite{Tanford1980} this is valid due to the condensed nature of the monolayer at this surface concentration resulting in the extended, staggered conformation of the chain being likely. In total for each co-refinement of two XRR measurements, there were ten degrees of freedom in the fitting process.

\begin{table}[h]
	\small
	\caption{\ The invariant parameters within the chemically-sensible model. 
	$^a$Values obtained from the Tanford formula where the carbon chains are assumed to be fully extended.\cite{Tanford1980} $^b$Values extracted from Sanchez-Fernandez \emph{et al.}\cite{Sanchez-Fernandez2016}}
	\label{tab:invariant}
	\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}lllll}
		\hline
		Component & $b_t$/fm & $b_h$/fm & $t_t$/\AA & $\text{SLD}$/$\times10^{-6}$\AA$^{-2}$ \\
		\hline
		X-ray & & & & \\
		DLPC & 5073 & 4674 & 15.5$^a$ & -- \\
		DMPC & 5985 & 4674 & 18.0$^a$ & -- \\
		DPPC & 6897 & 4674 & 20.5$^a$ & -- \\
		DMPG & 5985 & 4731 & 18.0$^a$ & --\\
		Air & -- & -- & -- & 0\\
		DES & -- & -- & -- & 10.8$^b$ \\
		\hline
		Neutron & & & & \\
		d$_{54}$-DMPC & 5329.8 & 602.7 & 18.0$^a$ & -- \\
		d$_{62}$-DPPC & 6129.2 & 602.7 & 20.5$^a$ & -- \\
		h-DES & -- & -- & -- & 0.43$^b$  \\
		hd-DES & -- & -- & -- & 3.15$^b$ \\
		\hline
	\end{tabular*}
\end{table}

In the second step, the determined head group and tail group component volumes were used in the refinement of the custom model against the NR measurements. Due to the lack of contrast present between the hydrogenous PC head group and the solvent, it was necessary to constrain the thickness of the head layer based on that determined from the lowest surface concentration XRR measurement. Since the NR measurements did not make use of a Langmuir trough sample environment, the $\phi_t$ parameter was allowed to have non-zero values. These values can be rationalised as a reduction of the compression of the tail components. To ensure that the number denisty of the head groups and pairs of tail groups was constant, the following constraint was applied to the value of $\phi_t$, 
\begin{equation}
\label{equ:phit}
\phi_t =  1 - \bigg(\frac{d_hV_t(1-\phi_h)}{V_hd_t}\bigg).
\end{equation}
Table \ref{tab:invariant} gives the details of the scattering length and SLDs used as invariant parameters in the custom model, while the tail chain lengths were the same as for the XRR.

In both cases, the refinement of the custom model to the experimental data involved transforming the reflectometry calculated from the model and the data into $Rq^4$ such that the artifact of the Fresnel decay was removed, before using the differential evoulation method available to refnx from the scipy library,\cite{Jones2001} to find the parameters that gave the best fit to the data. The parameter space was then probed in a Bayesian fashion using the MCMC method available through emcee, this allowed for an estimate of the PDF associated with each parameter. In the MCMC sampling, 200 walkers were used over 1000 iterations, following an equilibration of 200 iterations.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{figures/DLPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DPPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPG_all_data}
	\caption{The XRR profiles, SLD profiles and PDFs of the head component volume for each of the four lipids, the lower surface concentration is shown in blue while the higher in green; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. The XRR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by $5\times10^{-6}$ \AA$^{-2}$, for clarity. Figure files are available under MIT License.\cite{mccluskey_2018}}
	\label{fig:lipids}
\end{figure}
The custom model was co-refined across the two surface pressure measurements for each lipid. The resulting XRR profiles, associated SLD profiles and the PDF for the head group component volumes are shown in Figure \ref{fig:lipids}. Table \ref{tab:liptab} gives details of all varying parameters for each system, these are given with asymmetric uncertainies that correspond to a 95 \% confidence interval of the PDF, the full PDF plots can be found in the ESI. 
\begin{table*}
	\small
	\caption{\ The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the XRR models. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}. The values of $d_h$ were obtained from the appropriate use of Eqn. \ref{equ:ht}}
	\label{tab:liptab}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllllll}
		\hline
		Lipid & DLPC & & DMPC & & DPPC & & DMPG & \\
		Surface Pressure/mNm${-1}$ & 20 & 30 & 20 & 30 & 20 & 30 & 20 & 30 \\
		\hline
		$\theta_t$/$^\circ$ & \input{../output/dlpc/angle4.txt} & \input{../output/dlpc/angle5.txt} & \input{../output/dmpc/angle4.txt} & \input{../output/dmpc/angle5.txt} & \input{../output/dppc/angle4.txt} & \input{../output/dppc/angle5.txt} & \input{../output/dmpg/angle4.txt} & \input{../output/dmpg/angle5.txt} \\
